---
title: "Emotional 0-back model analysis"
author: "Maarten van der Velde"
date: "January 2018"
output: 
  html_notebook:
    toc: TRUE
    highlight: tango
    smart: no
  html_document:
    toc: TRUE
    highlight: tango
    smart: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

## Setup

```{r}
library(tidyverse)
library(forcats)
library(afex)
library(lsmeans)
library(tikzDevice)
options("tikzDocumentDeclaration" = "\\documentclass[12pt]{article}\n") # Default is 10pt.

use_tikz = FALSE # set to TRUE to save .tex versions of the plots

data_path <- "../data/0back/"
fig_path <- "../fig/"


# Plot groups
data_names = c("control", "depressed")
model_names = c("control model", "depressed model")
plot_groups = c(rbind(data_names, model_names))

# Plot theming
plot_theme <- theme_bw()

# Custom colour scale for plots
group_colours <- c("#5f91e2", "#5fc5e2", "#c4912d", "#e2ba5f")
names(group_colours) <- plot_groups
fill_scale <- scale_fill_manual(name = "", values = group_colours)

# Functions to calculate model fit to data
rsq <- function(model, data) {
  as_tibble(cbind(c("pearson", "kendall", "spearman"),
  c(cor(model, data, method = "pearson")^2,
  cor(model, data, method = "kendall")^2,
  cor(model, data, method = "spearman")^2)))
}

rmse <- function(model, data) {
  sqrt(mean((model - data)^2))
}

```



## Reading in the data

### Empirical data from Levens & Gotlib (2010)
```{r}
lgdat <- read.csv(paste0(data_path, "lgdat_0back.csv"))

lgdat$type <- factor(lgdat$type, levels=rev(levels(lgdat$type)))
lgdat$response <- factor(lgdat$response, levels = rev(levels(lgdat$response)))
```

### Model data
```{r}
control_dir <- paste0(data_path, "control50runs")
depressed_dir <- paste0(data_path, "depressed50runs")

beh_files <- c()
beh_files[1] <- tail(list.files(path = control_dir, pattern="beh.csv", full.names = TRUE),1)
beh_files[2] <- tail(list.files(path = depressed_dir, pattern="beh.csv", full.names = TRUE),1)

behdat_full <- data.frame()
for (i in 1:length(beh_files)) {
  behdat_full <- rbind(behdat_full, read.csv(beh_files[i], header=TRUE,sep=","))
}

behdat <- behdat_full %>%
  mutate(type = as.factor(ifelse(model == levels(model)[1], model_names[1], model_names[2]))) %>%
  mutate(id = as.factor(paste0(type, participant))) %>%
  mutate(expected_response = as.factor(ifelse(stimulus == target, "same", "diff")),
         accuracy = as.numeric(outcome == "correct"),
         responded = as.numeric(outcome %in% c("correct", "wrong"))) %>%
  select(-model, task_rep)


behdat_filtered <- behdat %>%
  filter(response != "none") %>% # Filter out non-responses
  filter(rt <= 2.0) %>% # Responses after 2s are too late
  filter(rt > 0) %>% # Response times of 0s indicate missed trials
  
  # Filter out trials with extreme RTs (outside 2.5 SDs of each participant's mean RT)
  group_by(type, participant) %>%
  mutate(rt_mean = mean(rt), rt_sd = sd(rt)) %>%
  filter(rt >= rt_mean - 2.5 * rt_sd, rt <= rt_mean + 2.5 * rt_sd) %>%
  droplevels()
  
```


## Accuracy

```{r}
accdat_model <- behdat_filtered %>%
  group_by(participant, type, stimulus, expected_response) %>%
  summarise(acc = mean(accuracy)) %>%
  group_by(type, stimulus, expected_response) %>%
  summarise(acc_mean = mean(acc), acc_sd = sd(acc))
  

accdat_lg <- lgdat %>%
  rename(stimulus = expression, expected_response = response) %>%
  mutate(acc_sd = NA) %>%
  select(stimulus, expected_response, acc_mean, acc_sd, type)


accdat_all <- rbind(accdat_lg, data.frame(accdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))


if (use_tikz) {
  tikz(file = paste0(fig_path, "0backAccuracyByResponse.tex"), width = 6, height = 3)
}

p <- ggplot(accdat_all, aes(x = expected_response, y = acc_mean, group = type, fill = type)) +
  facet_grid(~stimulus, labeller = labeller(stimulus = c(happy = "stimulus: happy", neutral = "stimulus: neutral", sad = "stimulus: sad"))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=acc_mean-acc_sd, ymax=acc_mean+acc_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Expected response", y = "Mean accuracy") +
  fill_scale + 
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}
```

### ANOVA & post-hoc contrasts
```{r}
# https://cran.r-project.org/web/packages/afex/vignettes/afex_anova_example.html
acc_aov <- aov_ez(id = "id", dv = "accuracy", data = behdat_filtered, between = "type", within = c("stimulus", "expected_response"), fun_aggregate = mean)
acc_aov

emmeans(acc_aov, ~ expected_response) %>%
  pairs(adjust = "bonf")

emmeans(acc_aov, ~ stimulus) %>%
  pairs(adjust = "bonf")

emmeans(acc_aov, ~ expected_response:stimulus) %>%
  pairs(adjust = "bonf")
```

### Fit to data
```{r}

acc_model <- accdat_model %>%
  arrange(type, stimulus, expected_response) %>%
  pull(acc_mean)

acc_data <- accdat_lg %>%
  arrange(type, stimulus, expected_response) %>%
  pull(acc_mean)

rsq(acc_model, acc_data)
rmse(acc_model, acc_data)
```


## Response rate
```{r}
respdat_model <- behdat %>%
  group_by(participant, type, target, expected_response, responded) %>%
  tally() %>%
  group_by(participant, type, target, expected_response) %>%
  mutate(freq = n/sum(n)) %>% # Represent as fraction
  filter(responded == TRUE) %>%
  select(-responded, -n) %>%
  group_by(type, target, expected_response) %>%
  summarise(rr_mean = mean(freq), rr_sd = sd(freq)) %>%
  rename(expression = target)

respdat_lg <- lgdat %>%
  mutate(rr_sd = NA) %>%
  rename(expected_response = response) %>%
  select(expression, expected_response, rr_mean, rr_sd, type)


respdat_all <- rbind(respdat_lg, data.frame(respdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))

if (use_tikz) {
  tikz(file = paste0(fig_path, "0backResponseRate.tex"), width = 6, height = 3)
}
p <- ggplot(respdat_all, aes(x = expected_response, y = rr_mean, group = type, fill = type)) +
  facet_grid(~expression, labeller = labeller(expression = c(happy = "target: happy", neutral = "target: neutral", sad = "target: sad"))) +
  geom_bar(stat = "identity", position=position_dodge(width=0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=rr_mean-rr_sd, ymax=rr_mean+rr_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Expected response", y = "Response rate") +
  fill_scale + 
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}
```

### ANOVA
```{r}
rr_aov <- aov_ez(id = "id", dv = "responded", data = behdat, between = "type", within = c("stimulus", "expected_response"), fun_aggregate = mean)
rr_aov

emmeans(rr_aov, ~ type) %>%
  pairs(adjust = "bonf")

emmeans(rr_aov, ~ expected_response) %>%
  pairs(adjust = "bonf")
```

### Fit to data
```{r}
rr_model <- respdat_model %>%
  arrange(type, expression, expected_response) %>%
  pull(rr_mean)

rr_data <- respdat_lg %>%
  arrange(type, expression, expected_response) %>%
  pull(rr_mean)

rsq(rr_model, rr_data)
rmse(rr_model, rr_data)
```


## Response time
```{r}
rtdat_model <- behdat_filtered %>%
  filter(outcome == "correct") %>%
  group_by(participant, type, stimulus, expected_response) %>%
  summarise(rt = mean(rt)) %>%
  group_by(type, stimulus, expected_response) %>%
  summarise(rt_mean = mean(rt), rt_sd = sd(rt)) %>%
  rename(expression = stimulus)

rtdat_lg <- lgdat %>%
  rename(expected_response = response) %>%
  select(expression, expected_response, rt_mean, rt_sd, type)

rtdat_all <- rbind(rtdat_lg, data.frame(rtdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))

if (use_tikz) {
  tikz(file = paste0(fig_path, "0backResponseTime.tex"), width = 6, height = 3)
}

p <- ggplot(rtdat_all, aes(x = expected_response, y = rt_mean, group = type, fill= type)) +
  facet_grid(~expression, labeller = labeller(expression = c(happy = "stimulus: happy", neutral = "stimulus: neutral", sad = "stimulus: sad"))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=rt_mean-rt_sd, ymax=rt_mean+rt_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Response", y = "Mean RT (s)") +
  fill_scale +
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}
```

### ANOVA
```{r}
rt_aov <- aov_ez(id = "id", dv = "rt", data = behdat_filtered, between = "type", within = c("stimulus", "expected_response"), fun_aggregate = mean)
rt_aov

emmeans(rt_aov, ~ type) %>%
  pairs(adjust = "bonf")

emmeans(rt_aov, ~ stimulus) %>%
  pairs(adjust = "bonf")

emmeans(rt_aov, ~ expected_response) %>%
  pairs(adjust = "bonf")

emmeans(rt_aov, ~ stimulus:expected_response) %>%
  pairs(adjust = "bonf")
```


## z-transformed response time
```{r}
behdat_filtered_z_rt <- behdat_filtered %>%
  select(-rt_mean, -rt_sd) %>%
  filter(outcome == "correct") %>%
  group_by(participant, type) %>%
  mutate(overall_rt = mean(rt)) %>%
  group_by(participant, type, stimulus, expected_response, overall_rt) %>%
  mutate(condition_rt = mean(rt), condition_rt_sd = sd(rt)) %>%
  mutate(z_rt = (condition_rt - overall_rt) / condition_rt_sd)


z_rtdat_model <- behdat_filtered_z_rt %>%
  group_by(type, stimulus, expected_response) %>%
  summarise(z_rt_mean = mean(z_rt), z_rt_sd = sd(z_rt)) %>%
  rename(expression = stimulus)


z_rtdat_lg <- lgdat %>%
  rename(expected_response = response) %>%
  select(expression, expected_response, z_rt_mean, z_rt_sd, type)


z_rtdat_all <- rbind(z_rtdat_lg, data.frame(z_rtdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))


if (use_tikz) {
  tikz(file = paste0(fig_path, "0backResponseTimeZ.tex"), width = 6, height = 3)
}

p <- ggplot(z_rtdat_all, aes(x = expected_response, y = z_rt_mean, group = type, fill= type)) +
  facet_grid(~expression, labeller = labeller(expression = c(happy = "stimulus: happy", neutral = "stimulus: neutral", sad = "stimulus: sad"))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=z_rt_mean-z_rt_sd, ymax=z_rt_mean+z_rt_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Response", y = "z-transformed mean RT") +
  fill_scale + 
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}
```

### ANOVA
```{r}
zrt_aov <- aov_ez(id = "id", dv = "z_rt", data = behdat_filtered_z_rt, between = "type", within = c("stimulus", "expected_response"), fun_aggregate = mean)
zrt_aov

emmeans(zrt_aov, ~ stimulus) %>%
  pairs(adjust = "bonf")

emmeans(zrt_aov, ~ expected_response) %>%
  pairs(adjust = "bonf")

emmeans(zrt_aov, ~ stimulus:expected_response) %>%
  pairs(adjust = "bonf")
```

### Fit to data
```{r}
zrt_model <- z_rtdat_model %>%
  arrange(type, expression, expected_response) %>%
  pull(z_rt_mean)

zrt_data <- z_rtdat_lg %>%
  arrange(type, expression, expected_response) %>%
  pull(z_rt_mean)

rsq(zrt_model, zrt_data)
rmse(zrt_model, zrt_data)
```



## Operator analysis
```{r}
op_files <- c()
op_files[1] <- tail(list.files(path = control_dir, pattern="ops.csv", full.names = TRUE),1)
op_files[2] <- tail(list.files(path = depressed_dir, pattern="ops.csv", full.names = TRUE),1)

opdat_full <- data.frame()
for (i in 1:length(op_files)) {
  opdat_full <- rbind(opdat_full, read.csv(op_files[i], header=TRUE,sep=","))
}

opdat <- opdat_full %>%
  select(-task_rep) %>%
  mutate(type = ifelse(model == levels(model)[1], model_names[1], model_names[2])) %>%
  mutate(on_task = as.logical(on_task), success = as.logical(success)) %>%
  select(-model)
```

### Frequency of mind-wandering: proportion of operators
```{r}
mw_freq <- opdat %>%
  group_by(type, participant, block, trial, on_task) %>%
  count(on_task) %>%
  group_by(type, participant, block, trial) %>%
  mutate(freq = n/sum(n)) %>%
  ungroup() %>%
  filter(on_task == FALSE) %>%
  mutate(freq = if_else(is.na(freq), 0, freq))


mw_share_of_ops <- mw_freq %>%
  group_by(type) %>%
  summarise(mw_mean = mean(freq), mw_sd = sd(freq))

mw_share_of_ops
```

### Frequency of mind-wandering: proportion of trials with >50% mind-wandering
```{r}
mw_dominant <- mw_freq %>%
  mutate(mw_trial = freq >= 0.5) %>%
  group_by(type) %>%
  count(mw_trial) %>%
  mutate(freq = nn/sum(nn))

mw_dominant
```


### Operator frequency
```{r}
op_freq <- opdat %>%
  filter(success == TRUE) %>%
  group_by(type, participant, operator, on_task) %>%
  count() %>%
  group_by(type, participant) %>%
  mutate(freq = n/sum(n)) %>%
  group_by(type, operator, on_task) %>%
  summarise(freq_mean = mean(freq), freq_sd = sd(freq))


p <- ggplot(op_freq, aes(x = operator, y = freq_mean, group = desc(type), fill = type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  coord_flip() +
  facet_grid(on_task ~ ., scales = "free_y", switch = "both", as.table = FALSE, labeller = labeller(on_task = c("TRUE" = "Task operators", "FALSE" = "Mind-wandering operators"))) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=freq_mean-freq_sd, ymax=freq_mean+freq_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Operator", y = "Frequency of use") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  fill_scale + 
  plot_theme

print(p)
```





