---
title: "Emotional 2-back model analysis"
author: "Maarten van der Velde"
date: "January 2018"
output:
  html_notebook:
    highlight: tango
    smart: no
    toc: yes
  html_document:
    highlight: tango
    smart: no
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

## Setup
```{r}
library(tidyverse)
library(stringr)
library(forcats)
library(afex)
library(lsmeans)
library(qgraph)
library(tikzDevice)
options("tikzDocumentDeclaration" = "\\documentclass[12pt]{article}\n") # Default is 10pt.

use_tikz = FALSE # set to TRUE to save .tex versions of the plots

data_path <- "../data/2back/"
fig_path <- "../fig/"

# Plot groups
data_names = c("control", "depressed")
model_names = c("control model", "depressed model")
plot_groups = c(rbind(data_names, model_names))

# Plot theming
plot_theme <- theme_bw()

# Custom colour scale for plots
group_colours <- c("#5f91e2", "#5fc5e2", "#c4912d", "#e2ba5f")
names(group_colours) <- plot_groups
fill_scale <- scale_fill_manual(name = "", values = group_colours)

```

## Reading in the data

### Empirical data from Levens & Gotlib (2010)
```{r}
lgdat <- read.csv(paste0(data_path, "lgdat_2back.csv"))
```

### Model data
```{r}
control_dir <- paste0(data_path, "20180121d")
depressed_dir <- paste0(data_path, "20180121e")

beh_files <- c()
beh_files[1] <- tail(list.files(path = control_dir, pattern="beh.csv", full.names = TRUE),1)
beh_files[2] <- tail(list.files(path = depressed_dir, pattern="beh.csv", full.names = TRUE),1)

behdat_full <- data.frame()
for (i in 1:length(beh_files)) {
  behdat_full <- rbind(behdat_full, read.csv(beh_files[i], header=TRUE,sep=","))
}

behdat <- behdat_full %>%
  mutate(type = ifelse(model == levels(model)[1], model_names[1], model_names[2])) %>%
  mutate(id = as.factor(paste0(type, participant))) %>%
  select(-model, -task_rep)

# Add outcome and condition of trial (see Levens and Gotlib (2010))
outcome <- rep("", nrow(behdat))
condition <- rep("", nrow(behdat))
valence <- behdat$stimulus

for (i in 1:nrow(behdat)) {
  
  # The first two trials of each block are discarded, no 2-back exists in either
  if (behdat[i,]$trial %% 55 %in% 1:2) {
    outcome[i] <- "none"
    condition[i] <- "none"
  } else {
    expected_answer <- ifelse(behdat[i,]$stimulus == behdat[i-2,]$stimulus, "same", "diff")
    outcome[i] <- ifelse(behdat[i,]$response == expected_answer, "correct", "wrong")
    
    if (expected_answer == "same") {
      # Match-set trial (same stimulus as two trials ago)
      condition[i] <- "match"
    } else {
      if (condition[i-1] == "match") {
        if (behdat[i-1,]$stimulus == behdat[i,]$stimulus) {
          # Perseverance-set trial (diff. than 2-back, but same as match-set in previous trial)
          condition[i] <- "pers"
        } else {
          # Break-set trial (diff. than 2-back, and diff. than match-set in previous trial)
          condition[i] <- "break"
          
          # In this case, the relevant valence is that of the broken set, not the current stimulus
          valence[i] <- valence[i-1] 
        }
      } else {
        condition[i] <- "noset"
      }
    }
  }
}

behdat$outcome <- as.factor(outcome)
behdat$condition <- as.factor(condition)
behdat$valence <- as.factor(valence)

behdat$accuracy <- as.numeric(behdat$outcome == "correct")
behdat$responded <- as.numeric(behdat$outcome %in% c("correct", "wrong") & between(behdat$rt, 0, 2.0) & behdat$response != "none")

behdat_filtered <- behdat %>%
  filter(response != "none") %>% # Filter out non-responses
  filter(rt <= 2.0) %>%   # Responses after 2s are too late
  filter(rt > 0) %>%      # Response times of 0s indicate missed trials
  filter(outcome != "none") %>% # Non-response trials at the start of each block

  # Filter out trials with extreme RTs (outside 2.5 SDs of each participant's mean RT)
  group_by(type, participant) %>%
  mutate(rt_mean = mean(rt), rt_sd = sd(rt)) %>%
  filter(rt >= rt_mean - 2.5 * rt_sd, rt <= rt_mean + 2.5 * rt_sd)

```


## Accuracy
```{r}
accdat_model <- behdat_filtered %>%
  group_by(participant, type, condition, valence) %>%
  summarise(acc = mean(accuracy)) %>%
  group_by(type, condition, valence) %>%
  summarise(acc_mean = mean(acc), acc_sd = sd(acc))

accdat_lg <- lgdat %>%
  rename(valence = expression) %>%
  mutate(acc_sd = NA) %>%
  select(valence, condition, acc_mean, acc_sd, type)

accdat_all <- rbind(accdat_lg, data.frame(accdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))


if (use_tikz) {
  tikz(file = paste0(fig_path, "2backAccuracyByCondition.tex"), width = 6, height = 3)
}

p <- ggplot(accdat_all, aes(x = valence, y = acc_mean, group = type, fill = type)) +
  facet_grid(~condition, labeller = labeller(condition = c("break" = "break-set", match = "match-set", noset = "no-set", pers = "perseverance-set"))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=acc_mean-acc_sd, ymax=acc_mean+acc_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Condition", y = "Mean accuracy") +
  fill_scale +
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}

accdat_all %>% group_by(type) %>% summarise(mean(acc_mean))
```

### ANOVA

```{r}
# Break-set condition
behdat_filtered_break <- behdat_filtered %>%
  filter(condition == "break")

acc_break_aov <- aov_ez(id = "id", dv = "accuracy", data = behdat_filtered_break, between = "type", within = "stimulus", fun_aggregate = mean)
acc_break_aov

# Match-set condition
behdat_filtered_match <- behdat_filtered %>%
  filter(condition == "match")

acc_match_aov <- aov_ez(id = "id", dv = "accuracy", data = behdat_filtered_match, between = "type", within = "stimulus", fun_aggregate = mean)
acc_match_aov

emmeans(acc_match_aov, ~ stimulus) %>%
  pairs(adjust = "bonf")

# No-set condition
behdat_filtered_noset <- behdat_filtered %>%
  filter(condition == "noset")

acc_noset_aov <- aov_ez(id = "id", dv = "accuracy", data = behdat_filtered_noset, between = "type", within = "stimulus", fun_aggregate = mean)
acc_noset_aov

# Perseverance-set condition
behdat_filtered_pers <- behdat_filtered %>%
  filter(condition == "pers")

acc_pers_aov <- aov_ez(id = "id", dv = "accuracy", data = behdat_filtered_pers, between = "type", within = "stimulus", fun_aggregate = mean)
acc_pers_aov
```

### Correlation with data
```{r}
acc_model <- accdat_model %>%
  arrange(type, condition, valence) %>%
  pull(acc_mean)

acc_data <- accdat_lg %>%
  arrange(type, condition, valence) %>%
  pull(acc_mean)

cor(acc_model, acc_data)

```

## Response rate
```{r}
respdat_model <- behdat %>%
  group_by(type, participant, condition, valence, responded) %>%
  tally() %>%
  group_by(type, participant, condition, valence) %>%
  mutate(freq = n/sum(n)) %>% # Represent as fraction
  filter(responded == TRUE) %>%
  select(-responded, -n) %>%
  group_by(type, condition, valence) %>%
  summarise(rr_mean = mean(freq), rr_sd = sd(freq))

respdat_lg <- lgdat %>%
  mutate(rr_sd = NA) %>%
  rename(valence = expression) %>%
  select(condition, valence, rr_mean, rr_sd, type)

respdat_all <- rbind(respdat_lg, data.frame(respdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))


if (use_tikz) {
  tikz(file = paste0(fig_path, "2backResponseRate.tex"), width = 6, height = 3)
}
p <- ggplot(respdat_all, aes(x = valence, y = rr_mean, group = type, fill = type)) +
  facet_grid(~condition, labeller = labeller(condition = c("break" = "break-set", match = "match-set", noset = "no-set", pers = "perseverance-set"))) +
  geom_bar(stat = "identity", position=position_dodge(width=0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=rr_mean-rr_sd, ymax=rr_mean+rr_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Condition", y = "Response rate") +
  fill_scale + 
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}

respdat_all %>% group_by(type) %>% summarise(mean(rr_mean))
```

### ANOVA

```{r}
# Break-set condition
behdat_break <- behdat %>%
  filter(condition == "break")

rr_break_aov <- aov_ez(id = "id", dv = "responded", data = behdat_break, between = "type", within = "stimulus", fun_aggregate = mean)
rr_break_aov

# Match-set condition
behdat_match <- behdat %>%
  filter(condition == "match")

rr_match_aov <- aov_ez(id = "id", dv = "responded", data = behdat_match, between = "type", within = "stimulus", fun_aggregate = mean)
rr_match_aov

# No-set condition
behdat_noset <- behdat %>%
  filter(condition == "noset")

rr_noset_aov <- aov_ez(id = "id", dv = "responded", data = behdat_noset, between = "type", within = "stimulus", fun_aggregate = mean)
rr_noset_aov

# Perseverance-set condition
behdat_pers<- behdat %>%
  filter(condition == "pers")

rr_pers_aov <- aov_ez(id = "id", dv = "responded", data = behdat_pers, between = "type", within = "stimulus", fun_aggregate = mean)
rr_pers_aov

```

### Correlation with data
```{r}
rr_model <- respdat_model %>%
  arrange(type, condition, valence) %>%
  pull(rr_mean)

rr_data <- respdat_lg %>%
  arrange(type, condition, valence) %>%
  pull(rr_mean)

cor(rr_model, rr_data)
```

## Response time
```{r}
rtdat_model <- behdat_filtered %>%
  filter(outcome == "correct") %>%
  group_by(type, participant, condition, valence) %>%
  summarise(rt = mean(rt)) %>%
  group_by(type, condition, valence) %>%
  summarise(rt_mean = mean(rt), rt_sd = sd(rt))

rtdat_lg <- lgdat %>%
  rename(valence = expression) %>%
  select(condition, valence, rt_mean, rt_sd, type)

rtdat_all <- rbind(rtdat_lg, data.frame(rtdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))

if (use_tikz) {
  tikz(file = paste0(fig_path, "2backResponseTime.tex"), width = 6, height = 3)
}

p <- ggplot(rtdat_all, aes(x = valence, y = rt_mean, group = type, fill= type)) +
  facet_grid(~condition, labeller = labeller(condition = c("break" = "break-set", match = "match-set", noset = "no-set", pers = "perseverance-set"))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=rt_mean-rt_sd, ymax=rt_mean+rt_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Condition", y = "Mean RT (s)") +
  fill_scale +
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}

rtdat_all %>% group_by(type) %>% summarise(mean(rt_mean))
```
### ANOVA
```{r}
# Break-set condition
rt_break_aov <- aov_ez(id = "id", dv = "rt", data = behdat_filtered_break, between = "type", within = "stimulus", fun_aggregate = mean)
rt_break_aov

# Match-set condition
rt_match_aov <- aov_ez(id = "id", dv = "rt", data = behdat_filtered_match, between = "type", within = "stimulus", fun_aggregate = mean)
rt_match_aov

# No-set condition
rt_noset_aov <- aov_ez(id = "id", dv = "rt", data = behdat_filtered_noset, between = "type", within = "stimulus", fun_aggregate = mean)
rt_noset_aov

# Perseverance-set condition
rt_pers_aov <- aov_ez(id = "id", dv = "rt", data = behdat_filtered_pers, between = "type", within = "stimulus", fun_aggregate = mean)
rt_pers_aov

```

### Correlation with data
```{r}
rt_model <- rtdat_model %>%
  arrange(type, condition, valence) %>%
  pull(rt_mean)

rt_data <- rtdat_lg %>%
  arrange(type, condition, valence) %>%
  pull(rt_mean)

cor(rt_model, rt_data)
```


## z-transformed response time
```{r}
behdat_filtered_z_rt <- behdat_filtered %>%
  select(-rt_mean, -rt_sd) %>%
  filter(outcome == "correct") %>%
  group_by(participant, type) %>%
  mutate(overall_rt = mean(rt)) %>%
  group_by(participant, type, condition, valence, overall_rt) %>%
  mutate(condition_rt = mean(rt), condition_rt_sd = sd(rt)) %>%
  mutate(z_rt = (condition_rt - overall_rt) / condition_rt_sd)

z_rtdat_model <- behdat_filtered_z_rt %>%
  group_by(type, condition, valence) %>%
  summarise(z_rt_mean = mean(z_rt), z_rt_sd = sd(z_rt))


z_rtdat_lg <- lgdat %>%
  rename(valence = expression) %>%
  select(valence, condition, z_rt_mean, z_rt_sd, type)

z_rtdat_all <- rbind(z_rtdat_lg, data.frame(z_rtdat_model)) %>%
  mutate(type = fct_relevel(type, plot_groups))

if (use_tikz) {
  tikz(file = paste0(fig_path, "2backResponseTimeZ.tex"), width = 6, height = 3)
}

p <- ggplot(z_rtdat_all, aes(x = valence, y = z_rt_mean, group = type, fill= type)) +
  facet_grid(~condition, labeller = labeller(condition = c("break" = "break-set", match = "match-set", noset = "no-set", pers = "perseverance-set"))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=z_rt_mean-z_rt_sd, ymax=z_rt_mean+z_rt_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Condition", y = "z-transformed mean RT") +
  fill_scale + 
  plot_theme

print(p)
if (use_tikz) {
  dev.off()
}

```

### ANOVA
```{r}
# Break-set condition
zrt_break <- behdat_filtered_z_rt %>%
  filter(condition == "break")

zrt_break_aov <- aov_ez(id = "id", dv = "z_rt", data = zrt_break, between = "type", within = "stimulus", fun_aggregate = mean)
zrt_break_aov

# Match-set condition
zrt_match <- behdat_filtered_z_rt %>%
  filter(condition == "match")

zrt_match_aov <- aov_ez(id = "id", dv = "rt", data = zrt_match, between = "type", within = "stimulus", fun_aggregate = mean)
zrt_match_aov

# No-set condition
zrt_noset <- behdat_filtered_z_rt %>%
  filter(condition == "noset")

zrt_noset_aov <- aov_ez(id = "id", dv = "rt", data = zrt_noset, between = "type", within = "stimulus", fun_aggregate = mean)
zrt_noset_aov

# Perseverance-set condition
zrt_pers <- behdat_filtered_z_rt %>%
  filter(condition == "pers")

zrt_pers_aov <- aov_ez(id = "id", dv = "rt", data = zrt_pers, between = "type", within = "stimulus", fun_aggregate = mean)
zrt_pers_aov
```



### Correlation with data
```{r}
zrt_model <- z_rtdat_model %>%
  arrange(type, condition, valence) %>%
  pull(z_rt_mean)

zrt_data <- z_rtdat_lg %>%
  arrange(type, condition, valence) %>%
  pull(z_rt_mean)

cor(zrt_model, zrt_data)
```

## Operator analysis
```{r}
op_files <- c()
op_files[1] <- tail(list.files(path = control_dir, pattern="ops.csv", full.names = TRUE),1)
op_files[2] <- tail(list.files(path = depressed_dir, pattern="ops.csv", full.names = TRUE),1)

opdat_full <- data.frame()
for (i in 1:length(op_files)) {
  opdat_full <- rbind(opdat_full, read.csv(op_files[i], header=TRUE,sep=","))
}

opdat <- opdat_full %>%
  select(-task_rep) %>%
  mutate(type = ifelse(model == levels(model)[1], model_names[1], model_names[2])) %>%
  mutate(on_task = as.logical(on_task), success = as.logical(success)) %>%
  select(-model)

```


### Frequency of mind-wandering: proportion of operators
```{r}
mw_freq <- opdat %>%
  group_by(type, participant, block, trial, on_task) %>%
  count(on_task) %>%
  group_by(type, participant, block, trial) %>%
  mutate(freq = n/sum(n)) %>%
  ungroup() %>%
  filter(on_task == FALSE) %>%
  mutate(freq = if_else(is.na(freq), 0, freq))


mw_share_of_ops <- mw_freq %>%
  group_by(type) %>%
  summarise(mw_mean = mean(freq), mw_sd = sd(freq))

mw_share_of_ops
```

### Frequency of mind-wandering: proportion of trials with >50% mind-wandering
```{r}
mw_dominant <- mw_freq %>%
  mutate(mw_trial = freq >= 0.5) %>%
  group_by(type) %>%
  count(mw_trial) %>%
  mutate(freq = nn/sum(nn))

mw_dominant
```

### Operator frequency
```{r}
op_freq <- opdat %>%
  filter(success == TRUE) %>%
  group_by(type, participant, operator, on_task) %>%
  count() %>%
  group_by(type, participant) %>%
  mutate(freq = n/sum(n)) %>%
  group_by(type, operator, on_task) %>%
  summarise(freq_mean = mean(freq), freq_sd = sd(freq))




ggplot(op_freq, aes(x = operator, y = freq_mean, group = desc(type), fill = type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  coord_flip() +
  facet_grid(on_task ~ ., scales = "free_y", switch = "both", as.table = FALSE, labeller = labeller(on_task = c("TRUE" = "Task operators", "FALSE" = "Mind-wandering operators"))) +
  scale_y_continuous() +
  geom_errorbar(aes(ymin=freq_mean-freq_sd, ymax=freq_mean+freq_sd), width=0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Operator", y = "Frequency of use") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  fill_scale + 
  plot_theme

```

## Transition probabilities between operators
```{r}
calc.trans.probs <- function(dat, variable) {
  
  dat <- droplevels(dat)
  
  vals <- sort(unique(dat[[variable]]))
  
  p <- matrix(nrow = length(vals), ncol = length(vals), 0)
  row.names(p) <- vals
  colnames(p) <- vals
  
  for (i in 1:(nrow(dat) - 1)) {
    # Go over items in sets of two
    # If they're from the same participant, count the transition
    if (dat$participant[i] == dat$participant[i+1] & dat$type[i] == dat$type[i+1]) {
      p[dat[[variable]][i], dat[[variable]][i+1]] <- p[dat[[variable]][i], dat[[variable]][i+1]] + 1
    }
  }
  
  # Convert counts to probability
  p <- p / rowSums(p)

  return(p)
}

plot.trans.probs.ops <- function(p, style = "circular") {
  # Plot transition probabilities between operators
  
  on_task = which(!str_detect(labels(p)[[1]], "wander|memory"))
  wandering = which(str_detect(labels(p)[[1]], "wander|memory"))
  groups = list("on task" = on_task, "wandering" = wandering)
  
  p[round(p,2) == 0] = NA # Remove zero-probabilities from plot
  
  qgraph(
    p, # transition probabilities, rounded to two decimal places
    groups = groups,
    layout = style,
    layout.control = 0.9,
    vsize = 5,
    esize = 5,
    #mar = c(8,7,8,7), # Margin size (bottom, left, top, right)
    labels = labels(p)[[1]],
    label.scale = FALSE,
    curveAll = TRUE,
    curveDefault = 1.05,
    edge.labels = TRUE, # Show numbers on arrows
    edge.label.cex = 1, # Size of edge labels
    border.width = 2, # Width of the borders of the nodes
    colFactor = 0.15, # controls degree of transparancy
    asize = 2, # Size of arrow heads
    maximum = 0.2 # Maximum transition probability, used for scaling
    #filetype = "pdf", # set to "pdf" or "tex" for standalone files
    #filename = paste0(gsub(" ", "_", graph_title))
  )
}



opdatcontrol <- opdat %>%
  mutate(on_task = if_else(on_task, "on task", "wandering")) %>%
  filter(type == "control model", success == TRUE)

opdatdepr <- opdat %>%
  mutate(on_task = if_else(on_task, "on task", "wandering")) %>%
  filter(type =="depressed model", success == TRUE) 
```

### Between on-task and off-task
```{r}
calc.trans.probs(opdatcontrol, "on_task") %>%
  plot.trans.probs.ops()

calc.trans.probs(opdatdepr, "on_task") %>%
  plot.trans.probs.ops()
```

### Between individual operators
```{r}
calc.trans.probs(opdatcontrol, "operator") %>%
  plot.trans.probs.ops()



calc.trans.probs(opdatdepr, "operator") %>%
  plot.trans.probs.ops()
```

## Retrieval analysis
```{r}
mem_files <- c()
mem_files[1] <- tail(list.files(path = control_dir, pattern="mems.csv", full.names = TRUE),1)
mem_files[2] <- tail(list.files(path = depressed_dir, pattern="mems.csv", full.names = TRUE),1)

memdatfull <- data.frame()
for (i in 1:length(mem_files)) {
  memdatfull <- rbind(memdatfull, read.csv(mem_files[i], header=TRUE,sep=","))
}

memdat <- memdatfull %>%
  mutate(group = ifelse(model == levels(model)[1], model_names[1], model_names[2])) %>%
  select(-model, -task_rep)


memdat %>%
  group_by(type, group) %>%
  summarise(count = n(), latency_mean = mean(latency))

```

